\chapter{Submitting a bug report}\label{sec:submitbug}

Whenever encountering that \TRACE fails while instrumenting an application or generating a trace-file, you may consider to submit a bug report to \texttt{tools@bsc.es}.
Before submitting a bug report, consider looking at the Frequently Asked Questions in Appendix \ref{sec:FAQ} because it may contain valuable information to address the failure you observe.

In any case, if you find that \TRACE fails and you want to submit a bug report, please collect as much as information possible to ease the bug-hunting process.
The information required depen whether the bug refers to a compilation or an execution issue.

\section{Reporting a compilation issue}

The following list of items are valuable when reporting a compilation problem:

\begin{itemize}
\item \TRACE version (this information appears in the first messages of the execution).
\item \TRACE configuration information, such as:
\begin{itemize}
  \item the configure output and the generated \texttt{config.log} file,
  \item versions of any additional libraries (PAPI, libunwind, DynInst, CUDA, OpenCL, libxml2, libdwarf, libelf, ...)
\end{itemize}
\item The compilation error itself as reported by invoking \texttt{make V=1}
\end{itemize}

\section{Reporting an execution issue}

The following list of items are valuable when reporting an execution problem:

\begin{itemize}
\item \TRACE version (this information appears in the first messages of the execution).
\item \TRACE configuration information, such as:
\begin{itemize}
  \item the configure output and the generated \texttt{config.log} file,
  \item versions of any additional libraries (PAPI, libunwind, DynInst, CUDA, OpenCL, libxml2, libdwarf, libelf, ...), and/or
  \item the \texttt{\$EXTRAE\_HOME}/etc/configured.sh output.
\end{itemize}
\item The result of the \texttt{make check} command after the \TRACE compilation process.
\item Does the application successfully executes with and without \TRACE?
\item Any valuable information from the system (type of processor, network, ...).
\item Which type of parallel programming paradigm does the application use: MPI, OpenMP, OmpSs, pthreads, ..., hybrid?
\item How do you execute the application? Which instrumentation library do you use?
\item The \TRACE configuration file used.
\item Any output generated by the application execution (in either the output or error channels).
\item If the execution generates a core dump, a backtrace of the dump using the \texttt{where} command of the \texttt{gdb} debugger.

\end{itemize}
